Galaxy Zoo: Morphological Classifications for Galaxies in HST Legacy Imaging

### 2016-01-15 writeup/gz_hubble_data.tex

\begin{itemize}
\item AEGIS has 1 orbit each (~2200 seconds) in F606W (V band) and F814W filter and has been dithered to 0.03 ''/pixel; the imaging covers $\sim710$~arcmin$^2$.

\item GOODS targeted 2 fields, GOODS-N and GOODS-S, imaging in 4 filters -- F435W (B), F606W (V), F775W (i), and F850LP (z). The mean exposure times vary, from 1000 - 2100 seconds. The images have been dithered to a pixel scale of 0.03 ''/pixel and covers at total area of ~320 arcmin$^2$ (160 arcmin$^2$ per field). The filters that \citet{gri12} uses for the colored images were F606W and F775W for GOODS-N and F606W and F850LP for GOODS-S.

\item COSMOS has 1 orbit (2028 seconds) in the F814W (I band) filter and has been dithered to 0.05 ''/pixel; it covers the largest area, ~1.8 deg$^2$.

\item GEMS has 1 orbit (2160 and 2286 seconds) in the F606W and F850LP filters, with a pixel scale of 0.03 ''/pixel; it covers ~800 arcmin$^2$
\end{itemize}

\begin{table*}
\caption{Summary of GZ Hubble Samples \label{gzh_numbers}}
\begin{tabular}{lclccc}
\hline\hline
Survey &  $t_{\rm exp}$ & Filters & Resolution & Area & $N_{\rm galaxies}$ \\
 & seconds & & \arcsec/pix & sq arcmin \\
\hline
AEGIS & 2200 & F606W (V) and F814W (I) & 0.03 & 710 \\
COSMOS & 2028 & F814W (I) & 0.05 & 6480\\
GEMS & 2160/2286 & F606W (V) and F850LP (z) & 0.03 & 800 \\
GOODS & 1000-2100s & F435W (B), F606W (V), F775W (i), F850LP (z) & 0.03 & 320 \\
\hline\hline
\end{tabular}
\end{table*}

\subsection{User Weighting}

@@

\section{Correcting for Redshift Dependent Classification Bias}

@@

In the GZH samples the redshift range is so large that we expect to find redshift evolution of the types and morphologies of galaxies which are seen. So the previous methods of correcting for redshift dependent bias will not work. In addition the effects of band shifting will change the images even more across these redshift ranges. Figure \ref{redshift} which illustrates one possible effect of losing features in spiral galaxies at high redshift).

%-----------------------------------------------------------------------------------------------------------------------------------
\begin{figure}
\begin{center}
\includegraphics[width=60mm]{figures/example_ferengi2.pdf}
\caption{Examples of an obvious spiral and obvious elliptical at $z=0$ which both look similar when redshifted to $z=0.8$. See below for the description of the method for redshifting.\label{exampleFERENGI}}
\end{center}
\end{figure}
 %-----------------------------------------------------------------------------------------------------------------------------------

In order to test the effects of redshift we made images of the same galaxy at a variety of redshifts using input images from the SDSS \citep{yor00,str02} and the \ferengi{} code \citep{bar08a} to produce images that match the observational properties of those observed in the HST surveys.

@@

We selected 288 unique galaxies from SDSS imaging to run through the \ferengi{} code.

The selection spanned a variety of galaxy morphologies (as indicated by GZ2 classifications), with a range of surface brightnesses, and also spanned the redshift range of SDSS targets (in $N_z = 4$ bins) in order to be optimised for different target minimum redshifts in HST imaging.

 The selection criteria for the different morphological categories is summarised in Table \ref{morphologies}. The surface brightness selection ($N_\mu = 3$) was (1) low: $\mu > 21.5 {\rm mag arsec}^2$;  (2) mid: $20.5 < \mu < 21.5 {\rm mag arsec}^2$; and (3) high: $\mu < 20.5 {\rm mag arsec}^2$.

@@

For each of the four ``target redshifts'' ($z = 0.3, 0.5, 0.8$ and $1.0$), the images were redshifted in $\Delta z = 0.1$ bins up to $z=1.0$. In addition different evolution models were assumed in \ferengi{}. This evolution is a crude mechanism that is suppose to mimic the brightness increase of galaxies with increasing redshift. The evolution mechanism allows us to make galaxies brighter as linear function of redshift, and that is it. It is an empirical addition to the magnitude of a galaxy of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, $e$ is the evolutionary correction in magnitudes.  (i.e. $e=-1$ essentially brightens the galaxy by 1~magnitude by $z=1$). We ran \ferengi{} for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in $\Delta e = 0.5$ steps. Depending on the target, or starting redshift we ran \ferengi{} for several final redshifts and different evolution models as summarized in Table~\ref{ferengivalues}.

\begin{table}
\caption{Summary of FERENGI runs \label{ferengivalues}}
\begin{tabular}{lcccccc}
\hline\hline
$z_{\rm target}$ & $N_{z {\rm bins}}$ & $N_{\rm evolution}$ & $e_{\rm max}$ & $N_{\rm runs}$ & $N_{\rm total}$\\
\hline
0.3              & 8                  & 7                   & $-3.0$        & 56             & 4031 \\
0.5              & 6                  & 4                   & $-1.5$        & 24             & 1728 \\
0.8              & 3                  & 3                   & $-1.0$        &  9             &  648 \\
1.0              & 1                  & 3                   & $-1.0$        &  3             &  216 \\
\hline\hline
\end{tabular}
\end{table}

Two series of example redshifted images are shown in Figure \ref{exampleFERENGI}, both for runs with no evolution ($e=0$).

@@

\subsection{Approach to debiasing morphological classifications}
\label{sec:zeta}

The approach used in GZH for correcting the weighted classifications for user bias rests on the assumption that the \emph{amount} of bias is a function of the apparent size and brightness of the image as seen on screen. This is controlled by two types of parameters: \textbf{intrinsic} properties of the galaxy itself, such as its physical diameter and luminosity, and \textbf{extrinsic} properties, such as the distance (redshift) of the galaxy and its relative orientation. There are likely other parameters that affect user accuracy, such as the proximity of close companions \citep[``distraction bias''; see][]{joh15} or those that vary by individual user. The combination of all such parameters forms a high-dimensional space, and we have insufficient test galaxies to measure the individual effects of each. Instead, we use just two parameters that are intended to capture the bulk of the change in bias (based on GZ1/GZ2): a galaxy's $r$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$; extrinsic).

The change in bias as a function of $\mu_r$ and $z$ is measured using the \ferengi{} images at all evolutionary corrected factors. We assume that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for a galaxy can be expressed as:

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/eye_of_sauron.pdf}
\caption{Surface brightness vs. redshift of 118,083 galaxies in the ACS Hubble sample. The white grid denotes the surface brightness and redshift range of the \ferengi{} images, subdivided in bins corresponding to fixed ranges used for analysis in Figure \ref{fig:p_vs_p}.}

@@

In Figure \ref{fig:p_vs_p} we examine the change in $p_{\rm features}$ for the \ferengi{} galaxies relative to their lowest simulated redshift. In this analysis, only galaxies whose lowest simulated redshift image was ($\rm z_{sim}=0.3$) were used (see Table \ref{ferengivalues}), and only those which had detectable surface brightness measurements in Sextractor; this includes 3,959 of the total 6,466 images. For each simulated redshift value $z$, and at fixed surface brightness $\mu$, we plot $p_{\rm features,z}$, the value measured at that simulated redshift, vs $p_{\rm features,z=0.3}$, the value measured for the same galaxy imaged at $z=0.3$.

@@

In Figure \ref{fig:p_vs_p} we see that the relationship between $p_{\rm features,z}$ and $p_{\rm features,z=0.3}$ is \emph{not} always single valued; hence we do not find it appropriate to correct galaxies that lie in these regions of surface brightness/redshift/$p_{\rm features}$ space. These regions tend to have low $p_{\rm features}$ values at high redshift, but a wide range of values at $z=0.3$. These regions contain two morphological types of galaxies: First are genuine ellipticals, which have low values of $p_{\rm features}$ at both high and low redshift. Second are disks whose features become washed out at high redshift; hence their $p_{\rm freatures}$ value at $z=0.3$ may be quite high, while the value observed at high redshift is very low. This effect is strongest at the highest redshift and the lowest surface brightnesses, where features become near impossible to discern from the images. From this information we conclude that, for the Hubble galaxies, we cannot distinguish between genuine ellipticals and disks if the measured value of $p_{\rm features}$ is too low for a given surface brightness and redshift, because the relationship between the measured value and the ``original'' ($z=0.3$) value is not single-valued.

Our criteria for determining whether a region of this space is single-valued, and therefore correctable, is as follows: For each surface brightness and redshift bin, we examine the distribution of $p_{\rm features,z=0.3}$ in four bins of $p_{\rm features,z}$. These sub-bins are marked by the horizontal lines within each larger, surface brightness/redshift bin in Figure \ref{fig:p_vs_p}. First, we compute the interquartile range (IQR) of the distribution of $p_{\rm features,z=0.3}$ in each sub-bin, which are indicated by the orange bars. For the region to be considered correctable, we require that the IQR of the distribution within these sub-bins is less than the sub-bin height. Regions which fit these criteria are uncouloured in the plot. If the IQR is greater than the bin height, the region is considered uncorrectable; these are marked in green. Any region with fewer than 5 points is considered to have ``not enough information (NEI)'', and are shown in purple.

@@

\begin{table}
\caption{Distribution of \ferengi{} images analysed in Figure \ref{fig:p_vs_p}. Correctable images had a single-valued relationship between their measured $p_{\rm features}$ values at high and low redshifts (white regions in Figure \ref{fig:p_vs_p}). Uncorrectable images had a non single-valued relationship (green regions). NEI images had undetermined relationships due to a lack of data ($N<5$) in their corresponding sub-bins (purple regions).   \label{ferengi_corrections}}
\begin{tabular}{lcc}
\hline \hline
				   & N     & \% \\
\hline 
Correctable                        & 2,057   & 52\% \\
Uncorrectable                      & 1,732   & 44\% \\
NEI                                & 170     &  4\%\\
Total                              & 3,959   & 100\% \\
\hline \hline
\end{tabular}
\end{table}

@@

We show here the {\bf preliminary} results of citizen scientist classifications of images of galaxies placed at artificial redshifts. Figure \ref{fig:ferengi_results_fake} shows the {\bf preliminary} range of change of vote fractions for the change of $p_{\rm features}$ with redshift, for galaxies with different vote fraction levels, three ranges of surface-brightness levels and 7~evolutionary corrections (this last bit is indicated by the colour).

@@

>>>>

\begin{itemize}
\item AEGIS has 1 orbit each ($\sim2200$~seconds) in the F606W (V-band) and F814W (I-band) filters and has been dithered to 0.03~\arcsec/pixel; the imaging covers $\sim710$~arcmin$^2$.

\item GOODS targeted 2 fields, GOODS-N and GOODS-S, imaging in 4 filters -- F435W (B), F606W (V), F775W (i), and F850LP (z). The mean exposure times vary, from 1000 -- 2100~seconds. The images have been dithered to a pixel scale of 0.03~\arcsec/pixel and covers a total area of $\sim320$~arcmin$^2$ (160~arcmin$^2$ per field). The filters that \citet{gri12} uses for the colored images were F606W and F775W for GOODS-N and F606W and F850LP for GOODS-S.

\item COSMOS has 1 orbit (2028~seconds) in the F814W (I-band) filter and has been dithered to 0.05~\arcsec/pixel; it covers the largest area, $\sim1.8$~deg$^2$.

\item GEMS has 1 orbit (2160 and 2286~seconds in the F606W and F850LP filters, respectively) with a pixel scale of 0.03~\arcsec/pixel; it covers $\sim800$~arcmin$^2$
\end{itemize}

% Note: galaxy counts remove the duplicate observations from gz_hst_table

\begin{table*}
\caption{Summary of Galaxy Zoo: Hubble imaging \label{gzh_numbers}}
\begin{tabular}{lclcrr}
\hline\hline
Survey &  $t_{\rm exp}$ & Filters & Resolution & Area & $N_{\rm galaxies}$ \\
 & [sec] & & [\arcsec/pix] & [arcmin$^2$] & \\
\hline
AEGIS                         & 2200        & F606W (V) and F814W (I)                     & 0.03 & 710   & 8157  \\
COSMOS                        & 2028        & F814W (I)                                   & 0.05 & 6480  & 88530 \\
GEMS                          & 2160$-$2286 & F606W (V) and F850LP (z)                    & 0.03 & 800   & 9143  \\
GOODS                         & 1000$-$2100 & F435W (B), F606W (V), F775W (i), F850LP (z) & 0.03 & 320   & 7336  \\
\hspace{10pt} \emph{GOODS-N}  & $-$         & $-$                                         & $-$  & $-$   & \emph{2551}  \\
\hspace{10pt} \emph{GOODS-S}  & $-$         & $-$                                         & $-$  & $-$   & \emph{4785}  \\
\hline
total                         & $-$         & $-$                                         & $-$  & 8310  & 113166  \\
\hline\hline
\end{tabular}
\end{table*}

\subsection{User weighting}

@@

\section{Correcting for redshift-dependent classification bias}

@@

In the GZH samples the redshift range is so large that we expect to find redshift evolution of the types and morphologies of galaxies which are seen. So the previous methods of correcting for redshift dependent bias will not work. In addition the effects of band shifting will change the images even more across these redshift ranges. %Figure~\ref{exampleFERENGI} illustrates some of the possible effects of losing features in spiral galaxies at high redshift.

%-----------------------------------------------------------------------------------------------------------------------------------
%\begin{figure}
%\begin{center}
%\includegraphics[width=60mm]{figures/example_ferengi2.pdf}
%\caption{Examples of an obvious spiral and obvious elliptical at $z=0$ which both look similar when redshifted to $z=0.8$. See below for the description of the method for redshifting.\label{exampleFERENGI}}
%\end{center}
%\end{figure}
 %-----------------------------------------------------------------------------------------------------------------------------------

In order to test the effects of redshift we made images of the same galaxy at a variety of redshifts using input images from the SDSS \citep{yor00,str02} and the \ferengi{} code \citep{bar08a} to produce images that match the observational properties of those observed in the HST surveys. These images were classified in the Galaxy Zoo interface using the same classification scheme as the original HST images.

@@

We selected 288 unique galaxies from SDSS imaging to run through the \ferengi{} code. The selection spanned a variety of galaxy morphologies (as selected by GZ2 classifications) and $r^\prime$-band surface brightnesses, and also spanned the redshift range of SDSS targets (in $N_z = 4$ bins) in order to be optimised for different target minimum redshifts in HST imaging.

The selection criteria for the different morphological categories is summarised in Table \ref{morphologies}. The surface brightness selection ($N_\mu = 3$) was (1) low: $\mu > 21.5$~\magarc;  (2) mid: $20.5 < \mu < 21.5$~\magarc; and (3) high: $\mu < 20.5$~\magarc. For each of the four ``target redshifts'' ($z = 0.3, 0.5, 0.8$ and $1.0$), the images were redshifted in $\Delta z = 0.1$ bins up to $z=1.0$.

@@

In addition to the physical parameters of the input images, the \ferengi{} output depends on assumptions of the global galaxy evolution model. This evolution is a crude mechanism that mimics the brightness increase of galaxies with increasing redshift (out to at least $z\sim1-2$). The effect on the redshifted images is simply an empirical addition to the magnitude of a galaxy of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, and $e$ is the evolutionary correction in magnitudes (i.e., $e=-1$ essentially brightens the galaxy by 1~magnitude by $z=1$). We ran \ferengi{} for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in increments of $\Delta e = 0.5$. Figure~\ref{exampleFERENGI} shows several examples of the effects of ``losing`` spiral/disc features with increasing redshift for two galaxies with $e=0$.

The final number of \ferengi{} images produced for each galaxy is a ultimately a function of galaxy's redshift, since the code cannot resample the images at better angular resolution than the original SDSS data, and the number of $e$ values selected. Table~\ref{ferengivalues} summarizes the total sample of redshifted images produced for GZ: Hubble.

\begin{table}
\caption{Summary of FERENGI artificial redshifting \label{ferengivalues}}
\begin{tabular}{lccccr}
\hline\hline
$z_{\rm target}$ & $N_{z {\rm bins}}$ & $N_{\rm evolution}$ & $e_{\rm max}$ & $N_{\rm galaxies}$ & $N_{\rm images}$\\
\hline
0.3              & 8                  & 7                   & $-3.0$        & 72             & 4032 \\
0.5              & 6                  & 4                   & $-1.5$        & 72             & 1728 \\
0.8              & 3                  & 3                   & $-1.0$        & 72             &  648 \\
1.0              & 1                  & 3                   & $-1.0$        & 72             &  216 \\
\hline\hline
\end{tabular}
\end{table}

@@

\subsection{Correcting GZH morphologies for classification bias}
\label{sec:zeta}

The approach used in GZH for correcting the weighted classifications for user bias rests on the assumption that the \emph{amount} of bias is a function of the apparent size and brightness of the image as seen on screen. This is controlled by two types of parameters: \textbf{intrinsic} properties of the galaxy itself, such as its physical diameter and luminosity, and \textbf{extrinsic} properties, such as the distance (redshift) of the galaxy and its relative orientation. There are likely other parameters that affect user accuracy, such as the proximity of close companions \citep[``distraction bias''; see][]{joh15} or those that vary by individual user. The combination of all such parameters forms a high-dimensional space, and we have insufficient test galaxies to measure the individual effects of each. Instead, we use just two parameters that are intended to capture the bulk of the change in bias (based on GZ1/GZ2): a galaxy's $r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$; extrinsic).

The change in bias as a function of $\mu_r$ and $z$ is measured using the \ferengi{} images over all the evolutionary corrected factors. We assume that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for a galaxy can be expressed as:

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/eye_of_sauron.pdf}
\caption{Surface brightness vs. redshift of 118,083 galaxies in the ACS Hubble sample. The white grid denotes the surface brightness and redshift range of the \ferengi{} images, subdivided in bins corresponding to fixed ranges used for analysis in Figure~\ref{fig:p_vs_p}.}

@@

In Figure~\ref{fig:p_vs_p} we examine the change in $p_{\rm features}$ for the \ferengi{} galaxies relative to their lowest simulated redshift. In this analysis, only galaxies whose lowest simulated redshift image was ($\rm z_{sim}=0.3$) were used (see Table \ref{ferengivalues}), and only those which had detectable surface brightness measurements in Sextractor; this includes 3,959 of the total 6,466 images. For each simulated redshift value $z$, and at fixed surface brightness $\mu$, we plot $p_{\rm features,z}$, the value measured at that simulated redshift, vs $p_{\rm features,z=0.3}$, the value measured for the same galaxy imaged at $z=0.3$.

@@

In Figure~\ref{fig:p_vs_p} we see that the relationship between $p_{\rm features,z}$ and $p_{\rm features,z=0.3}$ is \emph{not} always single valued; hence we do not find it appropriate to correct galaxies that lie in these regions of surface brightness/redshift/$p_{\rm features}$ space. These regions tend to have low $p_{\rm features}$ values at high redshift, but a wide range of values at $z=0.3$. These regions contain two morphological types of galaxies: First are genuine ellipticals, which have low values of $p_{\rm features}$ at both high and low redshift. Second are disks whose features become washed out at high redshift; hence their $p_{\rm freatures}$ value at $z=0.3$ may be quite high, while the value observed at high redshift is very low. This effect is strongest at the highest redshift and the lowest surface brightnesses, where features become near impossible to discern from the images. From this information we conclude that, for the Hubble galaxies, we cannot distinguish between genuine ellipticals and disks if the measured value of $p_{\rm features}$ is too low for a given surface brightness and redshift, because the relationship between the measured value and the ``original'' ($z=0.3$) value is not single-valued.

Our criteria for determining whether a region of this space is single-valued, and therefore correctable, is as follows: For each surface brightness and redshift bin, we examine the distribution of $p_{\rm features,z=0.3}$ in four bins of $p_{\rm features,z}$. These sub-bins are marked by the horizontal lines within each larger, surface brightness/redshift bin in Figure~\ref{fig:p_vs_p}. First, we compute the interquartile range (IQR) of the distribution of $p_{\rm features,z=0.3}$ in each sub-bin, which are indicated by the orange bars. For the region to be considered correctable, we require that the IQR of the distribution within these sub-bins is less than the sub-bin height. Regions which fit these criteria are uncouloured in the plot. If the IQR is greater than the bin height, the region is considered uncorrectable; these are marked in green. Any region with fewer than 5 points is considered to have ``not enough information (NEI)'', and are shown in purple.

@@

\begin{table}
\caption{Distribution of \ferengi{} images analysed in Figure~\ref{fig:p_vs_p}. Correctable images had a single-valued relationship between their measured $p_{\rm features}$ values at high and low redshifts (white regions in Figure~\ref{fig:p_vs_p}). Uncorrectable images had a non single-valued relationship (green regions). NEI images had undetermined relationships due to a lack of data ($N<5$) in their corresponding sub-bins (purple regions).   \label{ferengi_corrections}}
\begin{tabular}{lcc}
\hline \hline
				   & N     & \% \\
\hline 
Correctable                        & 2,057   & 52\% \\
Uncorrectable                      & 1,732   & 44\% \\
NEI                                & 170     &  4\%\\
Total                              & 3,959   & 100\% \\
\hline \hline
\end{tabular}
\end{table}

@@

We show here the {\bf preliminary} results of citizen scientist classifications of images of galaxies placed at artificial redshifts. Figure~\ref{fig:ferengi_results_fake} shows the {\bf preliminary} range of change of vote fractions for the change of $p_{\rm features}$ with redshift, for galaxies with different vote fraction levels, three ranges of surface-brightness levels and 7~evolutionary corrections (this last bit is indicated by the colour).

@@

\subsection{Duplicate images}

\subsection{SDSS images}
